---
output: github_document
---

```{r, echo=FALSE}

knitr::opts_chunk$set(
  fig.path = "man/figures/README-",
  out.width = "100%"
)

```


# ggcustom

`ggcustom` is an R package that provides custom themes and color scales for `ggplot2` visualizations, including the VM color palette and theme as one of the available style options.

## Installation

To install `ggcustom` directly from GitHub, use the `devtools` package:

```r
# Install devtools if not already installed
install.packages("devtools")

# Install ggcustom from GitHub
devtools::install_github("jhuovari/ggcustom")

```

#  Overview

The `ggcustom` package is designed to make it easy to apply consistent themes and color schemes to ggplot2 visualizations. It includes:

- Custom color palettes for FPB and VM.
- Discrete color and fill scales using those palettes.
- FPB and VM themes for ggplot2 plots

#  Usage

Once the package is installed, you can load it with:

```{r}

library(ggcustom)

```

Color Scales
The package includes custom color scales based on the VM color palette.

```{r}

library(ggplot2)
library(dplyr)

# Example dataset
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]

# Using the VM color scale
ggplot(dsamp, aes(carat, price, colour = clarity)) +
  geom_point() +
  scale_colour_vm()
  
```

You can also use the fill scale for bar plots or other filled geometries:

```{r}

ggplot(dsamp, aes(clarity, fill = clarity)) +
  geom_bar() +
  scale_fill_vm()
  
```

# VM Theme

The theme_vm function provides a customized ggplot2 theme based on theme_bw, with defaults that make use of the VM color palette.

```{r}

# Using the VM theme
p <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
p + theme_vm()

```

# Theme Finnish Productivity Board

```{r}

# Using the fpb theme
p <- txhousing |> 
  filter(city %in% c("Austin", "Houston", "Dallas", "Arlington")) |> 
  ggplot(aes(date, inventory, colour = city)) +
  geom_line() +
  labs(title = "Plot title", subtitle = "Subtitle", caption = "Source:TAMU real estate center")

p + scale_colour_manual(values = fpb_pal(4)) + theme_fpb()

```


### Setting Custom Theme and Palette with `set_gg()`

The `set_gg()` function allows you to set a custom theme and palette for your `ggplot2` visualizations. Below is an example using the `theme_vm()` and the `vm` palette. For VM there is also shortcut `set_vm()`.

```{r}


# Example dataset
data <- data.frame(
  category = rep(letters[1:4], each = 3),
  value = c(4, 3, 5, 2, 6, 7, 3, 5, 4, 6, 2, 7),
  group = rep(1:3, times = 4)
)

# Define a ggplot object
p <- ggplot(data, aes(x = category, y = value, colour = factor(group), group = group)) +
  geom_line() +
  geom_point()

# Before setting custom theme and palette
p

# Set the VM theme and VM palette
set_gg(theme_vm(), "vm")

# Plot with VM theme and VM color palette
p

# Optionally, reset to the original theme and scales
unset_gg()

# Verify that the original theme and scales are restored
p
```

# Additional Palettes and Themes

The package is structured to allow additional color palettes and themes, making it easy to expand with other organization-specific styles.

## Examples

```{r}
# Scatter plot with VM theme and color scale
ggplot(mtcars, aes(mpg, wt, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_colour_vm() +
  theme_vm()
```


```{r}
# Bar plot with VM fill scale
ggplot(mpg, aes(class, fill = class)) +
  geom_bar() +
  scale_fill_vm() +
  theme_vm()

```


## Theme helpers

The package includes a few small helpers to simplify adjusting common
[ggplot2](https://ggplot2.tidyverse.org) theme elements. They wrap
`\code{\link[ggplot2]{theme}}` calls with convenient defaults.

- `the_x45()` — rotate x-axis tick labels 45° (useful for long category names).
- `the_legend_bot()` — move the legend to the bottom.
- `the_title_blank()` — blank out selected titles (axis titles, plot title,
  legend title, caption). You can specify which elements to blank with short
  codes, e.g. `"xyt"` for x-axis, y-axis and plot title.

```{r, message=FALSE}
library(ggplot2)

p <- ggplot(mtcars, aes(factor(cyl), mpg, fill = factor(gear))) +
  geom_boxplot() +
  labs(title = "Example plot", x = "Cylinders", y = "MPG",
       caption = "Data: mtcars")

# Rotate x labels, legend at bottom, blank axis titles
p + the_x45() + the_legend_bot() + the_title_blank("xy")

```

# Contributing

Contributions are welcome! If you’d like to add new color palettes, themes, or functionality, please submit a pull request or open an issue.

# License

ggcustom is licensed under the MIT License.
