---
output: github_document
---

# ggcustom

`ggcustom` is an R package that provides custom themes and color scales for `ggplot2` visualizations, including the VM color palette and theme as one of the available style options.

## Installation

To install `ggcustom` directly from GitHub, use the `devtools` package:

```r
# Install devtools if not already installed
install.packages("devtools")

# Install ggcustom from GitHub
devtools::install_github("jhuovari/ggcustom")

```

#  Overview

The `ggcustom` package is designed to make it easy to apply consistent themes and color schemes to ggplot2 visualizations. It includes:

- Custom color palettes for FPB and VM.
- Discrete color and fill scales using those palettes.
- FPB and VM themes for ggplot2 plots

#  Usage

Once the package is installed, you can load it with:

```{r}

library(ggcustom)

```

Color Scales
The package includes custom color scales based on the VM color palette.

```{r}

library(ggplot2)

# Example dataset
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]

# Using the VM color scale
ggplot(dsamp, aes(carat, price, colour = clarity)) +
  geom_point() +
  scale_colour_vm()
  
```

You can also use the fill scale for bar plots or other filled geometries:

```{r}

ggplot(dsamp, aes(clarity, fill = clarity)) +
  geom_bar() +
  scale_fill_vm()
  
```

# VM Theme

The theme_vm function provides a customized ggplot2 theme based on theme_bw, with defaults that make use of the VM color palette.

```{r}

# Using the VM theme
p <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
p + theme_vm()

```

### Setting Custom Theme and Palette with `set_gg()`

The `set_gg()` function allows you to set a custom theme and palette for your `ggplot2` visualizations. Below is an example using the `theme_vm()` and the `vm` palette. For VM there is also shortcut `set_vm()`.

```{r}


# Example dataset
data <- data.frame(
  category = rep(letters[1:4], each = 3),
  value = c(4, 3, 5, 2, 6, 7, 3, 5, 4, 6, 2, 7),
  group = rep(1:3, times = 4)
)

# Define a ggplot object
p <- ggplot(data, aes(x = category, y = value, colour = factor(group), group = group)) +
  geom_line() +
  geom_point()

# Before setting custom theme and palette
p

# Set the VM theme and VM palette
set_gg(theme_vm(), "vm")

# Plot with VM theme and VM color palette
p

# Optionally, reset to the original theme and scales
unset_gg()

# Verify that the original theme and scales are restored
p
```

# Additional Palettes and Themes

The package is structured to allow additional color palettes and themes, making it easy to expand with other organization-specific styles.

# Examples

```{r}
# Scatter plot with VM theme and color scale
ggplot(mtcars, aes(mpg, wt, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_colour_vm() +
  theme_vm()
```


```{r}
# Bar plot with VM fill scale
ggplot(mpg, aes(class, fill = class)) +
  geom_bar() +
  scale_fill_vm() +
  theme_vm()

```




# Contributing

Contributions are welcome! If you’d like to add new color palettes, themes, or functionality, please submit a pull request or open an issue.

# License

ggcustom is licensed under the MIT License.
